MOM Estimator: Regression Form

### 2024-12-29 writeUps/regression.tex
update

Let $p_i=Pr(S_i=1|\bm{x}_i)$.
Then the original estimating equations were:
\begin{align*}
  Z_i=0:&\\
  &Y_i-\mu_{00}(1-p_i)+\mu_{01}p_i=Y_i-\mu_{00}-(\mu_{01}-\mu_{00})p_i\\
  p_i&Y_i-\mu_{00}p_i-(\mu_{01}-\mu_{00})p_i^2\\
  Z_i=1:&\\
        &Y_i-\mu_{10}-S_i(\mu_{11}-\mu_{10})\\
  S_i&Y_i-S_i\mu_{10}-S_i(\mu_{11}-\mu_{10})=S_iY_i-S_i\mu_{11}\\
\end{align*}

@@

\section{Missing $S$ in Treatment Group}
When $S$ is missing at random (conditional on $\bm{x}$) for some observations in the treatment group,
and $\mathbb{E}[Y|Z=1,S,p]=\mathbb{E}[Y|Z=1,S]$,
a similar mixed model will work for estimating $\mu_{10}$ and $\mu_{11}$.
\begin{align*}
  Z_i=1\text{ and }S_i\text{ Missing}:&\\
  &Y_i-\mu_{10}(1-p_i)+\mu_{11}p_i=Y_i-\mu_{10}-(\mu_{11}-\mu_{10})p_i\\
  p_i&Y_i-\mu_{10}p_i-(\mu_{11}-\mu_{10})p_i^2\\
  Z_i=1\text{ and }S_i\text{ Observed}:&\\
        &Y_i-\mu_{10}-S_i(\mu_{11}-\mu_{10})\\
  S_i&Y_i-S_i\mu_{10}-S_i(\mu_{11}-\mu_{10})=S_iY_i-S_i\mu_{11}\\
    Z_i=0:&\\
  &Y_i-\mu_{00}(1-p_i)+\mu_{01}p_i=Y_i-\mu_{00}-(\mu_{01}-\mu_{00})p_i\\
  p_i&Y_i-\mu_{00}p_i-(\mu_{01}-\mu_{00})p_i^2\\
\end{align*}

@@

So we're estimating principal effects in two stages.
First, fit the model
\begin{equation}\label{eq:psMod}
  Pr(S_{Ti}=1|\bm{\tilde{x}}_i)=f(\bm{\alpha}'\bm{\tilde{x}}_i)
\end{equation}
for some inverse link function $f(\cdot)$, where $\bm{\tilde{x}}_i=[1,\bm{x}_i]$, using (observed) values from the treatment group, and estimating $\hat{\alpha}$.
Then let
\begin{equation}\label{eq:ps}
  \hat{p}_i=f(\bm{\hat{\alpha}}'\bm{\tilde{x}}_i)
\end{equation}
for all subjects in the experiment.

Finally, fit model
\begin{equation}\label{eq:regression}
  Y_i=\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}^t\bm{x}_i+\epsilon_i
\end{equation}
to estimate $\bm{\beta}$ and hence principal effects.
[Note that \eqref{eq:regression} is just \eqref{eq:regression0} with $\bm{\gamma}^t\bm{x}_i$ added on the end].
Let $\bm{\beta}=[\beta_0,\beta_1,\beta_2,\beta_3,\bm{\gamma}^t]^t$.

The standard error matrix for $\bm{\hat{\beta}}$ can be estimated as:
\begin{equation}\label{eq:sandwich}
  \widehat{var}(\bm{\hat{\beta}})=A^{-1}BA^{-t}
\end{equation}
Where
\begin{equation}\label{eq:Amat}
  A=\sum_i\frac{\partial}{\partial [\alpha,\beta]^t} \Psi(Y_i;\alpha,\beta,\bm{x}_i)
\end{equation}
and
\begin{equation}\label{eq:Bmat}
  B=\sum_i \Psi(Y_i;\alpha,\beta,\bm{x}_i)\Psi(Y_i;\alpha,\beta,\bm{x}_i)^t
\end{equation}
where $\Psi(\cdot)$ is the estimating equations.

@@

\subsection{$A_{21}$ Matrix}
The estimating equations for the regression \eqref{eq:regression} are
\begin{equation}\label{eq:eeOLS}
  \psi(Y_i,\bm{x}_i,\beta,\alpha)=X_iY_i-X_iX_i^t\beta
\end{equation}
Where $X_i=[1,r_i,Z_i,r_iZ_i,\bm{x}_i^t]^t$.
These depend on $\alpha$ when $r_i=p_i$, i.e. when $Z_i=0$ or $S_i$ is missing.
Then note that, following \eqref{eq:ps}, and letting $\eta_i=\alpha^t\bm{\tilde{x}}_i$
\begin{equation}\label{eq:derivP}
  \frac{\partial p_i}{\partial \alpha^t}=f'(\eta_i)\frac{\partial \eta_i}{\partial \alpha^t}=f'(\eta_i)\bm{\tilde{x}}_i^t
\end{equation}
and that
\begin{equation}
  \frac{\partial p_i^2}{\partial \alpha^t}=2p\frac{\partial p_i}{\partial \alpha^t}=2f(\eta)f'(\eta_i)\bm{\tilde{x}}_i^t
\end{equation}

Then if $r_i=p_i$,
\begin{align*}
  \frac{\partial}{\partial \alpha^t}&\psi(Y_i,\bm{x}_i,\beta,\alpha)=\\
  &\left[\begin{array}{c}
          -\beta_1\frac{\partial p_i}{\partial \alpha^t}\\
          (Y_i-\beta_0-\beta_2Z_i-\bm{\gamma}^t\bm{x}_{i})\partial p/\partial\alpha^t-(\beta_1+\beta_3Z_i)\partial p^2/\partial \alpha^t\\
          -\beta_1\frac{\partial p_i}{\partial \alpha^t}Z_i\\
          Z_i\left[(Y_i-\beta_0-\beta_2Z_i-\bm{\gamma}^t\bm{x}_{i})\partial p/\partial\alpha^t-(\beta_1+\beta_3Z_i)\partial p^2/\partial \alpha^t\right]\\
          -\beta_1\frac{\partial p_i}{\partial \alpha^t}\bm{x}_i
        \end{array}\right]\\
  =& \left[\begin{array}{c}
          -\beta_1f'(\eta_i)\bm{\tilde{x}}_i^t\\
          \left[Y_i-X_i^t\beta-p_i(\beta_1+\beta_3Z_i)\right]f'(\eta_i)\bm{\tilde{x}}_i\\
          -Z_i\beta_1f'(\eta_i)\bm{\tilde{x}}_i^t\\
          Z_i\left[Y_i-X_i^t\beta-p_i(\beta_1+\beta_3)\right]f'(\eta_i)\bm{\tilde{x}}_i\\
          -\beta_1f'(\eta_i)\bm{x}_i\bm{x}_i^t
           \end{array}\right]
\end{align*}
Note that unless there is missing $S$ in the treatment group, $Z_i=0$ whenever $r_i=p_i$, so the 3rd and 4th lines of the matrix are all 0.

If $r_i=S_i$, $\frac{\partial}{\partial \alpha^t}\text{ \& }\psi(Y_i,\bm{x}_i,\beta,\alpha)=0$.

>>>>

Let $p_i=Pr(S_i=1|\bm{w}_i)$, where $\bm{w}_i$ is a vector of covariates that (putatively) predict $S_i$.
(Really this should be $S_{Ti}$, cuz we're doing principal stratification; we'll drop the $T$ subscript for clarity.)

Then the original estimating equations were:
\begin{align*}
  Z_i=0:&\\
  &Y_i-\mu_{00}(1-p_i)+\mu_{01}p_i=Y_i-\mu_{00}-(\mu_{01}-\mu_{00})p_i\\
  p_i&Y_i-\mu_{00}p_i-(\mu_{01}-\mu_{00})p_i^2\\
  Z_i=1:&\\
        &Y_i-\mu_{10}-S_i(\mu_{11}-\mu_{10})\\
  S_i&Y_i-S_i\mu_{10}-S_i(\mu_{11}-\mu_{10})=S_iY_i-S_i\mu_{11}\\
\end{align*}

@@

\section{Missing $S$ in Treatment Group}
When $S$ is missing at random (conditional on $\bm{w}$) for some observations in the treatment group,
and $\mathbb{E}[Y|Z=1,S,p]=\mathbb{E}[Y|Z=1,S]$,
a similar mixed model will work for estimating $\mu_{10}$ and $\mu_{11}$.
\begin{align*}
  Z_i=1\text{ and }S_i\text{ Missing}:&\\
  &Y_i-\mu_{10}(1-p_i)+\mu_{11}p_i=Y_i-\mu_{10}-(\mu_{11}-\mu_{10})p_i\\
  p_i&Y_i-\mu_{10}p_i-(\mu_{11}-\mu_{10})p_i^2\\
  Z_i=1\text{ and }S_i\text{ Observed}:&\\
        &Y_i-\mu_{10}-S_i(\mu_{11}-\mu_{10})\\
  S_i&Y_i-S_i\mu_{10}-S_i(\mu_{11}-\mu_{10})=S_iY_i-S_i\mu_{11}\\
    Z_i=0:&\\
  &Y_i-\mu_{00}(1-p_i)+\mu_{01}p_i=Y_i-\mu_{00}-(\mu_{01}-\mu_{00})p_i\\
  p_i&Y_i-\mu_{00}p_i-(\mu_{01}-\mu_{00})p_i^2\\
\end{align*}

@@

So we're estimating principal effects in two stages.
First, fit the model
\begin{equation}\label{eq:psMod}
  Pr(S_{Ti}=1|\bm{\tilde{w}}_i)=f(\bm{\alpha}'\bm{\tilde{w}}_i)
\end{equation}
for some inverse link function $f(\cdot)$, where $\bm{\tilde{x}}_i=[1,\bm{w}_i]$, using (observed) values from the treatment group, and estimating $\hat{\alpha}$.
Then let
\begin{equation}\label{eq:ps}
  \hat{p}_i=f(\bm{\hat{\alpha}}'\bm{\tilde{w}}_i)
\end{equation}
for all subjects in the experiment.

Finally, fit model
\begin{equation}\label{eq:regression}
  Y_i=\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}^t\bm{x}_i+\epsilon_i
\end{equation}
to estimate $\bm{\beta}$ and hence principal effects, where $\bm{x}_i$ is a set of covariates predictive of $Y$ within principal strata.
[Note that \eqref{eq:regression} is just \eqref{eq:regression0} with $\bm{\gamma}^t\bm{x}_i$ added on the end].
$\bm{x}_i$ should, generally, contain $\bm{w}_i$; the two may also be equivalent.
Let $\bm{\beta}=[\beta_0,\beta_1,\beta_2,\beta_3,\bm{\gamma}^t]^t$.

The standard error matrix for $\bm{\hat{\beta}}$ can be estimated as:
\begin{equation}\label{eq:sandwich}
  \widehat{var}(\bm{\hat{\beta}})=A^{-1}BA^{-t}
\end{equation}
Where
\begin{equation}\label{eq:Amat}
  A=\sum_i\frac{\partial}{\partial [\alpha,\beta]^t} \Psi(Y_i;\alpha,\beta,\bm{x}_i,\bm{w}_i)
\end{equation}
and
\begin{equation}\label{eq:Bmat}
  B=\sum_i \Psi(Y_i;\alpha,\beta,\bm{x}_i)\Psi(Y_i;\alpha,\beta,\bm{x}_i,\bm{w}_i)^t
\end{equation}
where $\Psi(\cdot)$ is the estimating equations.

@@

\subsection{$A_{21}$ Matrix}
The estimating equations for the regression \eqref{eq:regression} are
\begin{equation}\label{eq:eeOLS}
  \psi(Y_i,\bm{x}_i,\bm{w}_i,\beta,\alpha)=X_iY_i-X_iX_i^t\beta
\end{equation}
Where $X_i=[1,r_i,Z_i,r_iZ_i,\bm{x}_i^t]^t$.
In other words,
\begin{align*}
  \psi(Y_i,&\bm{x}_i,\bm{w}_i,\beta,\alpha)=\\
  &\left\{
  \begin{array}{l}
    Y_i-\left(\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}'\bm{x}_i\right)\\
    r_iY_i-r_i\left(\beta_0+\beta_2Z_i+\bm{\gamma}'\bm{x}_i\right)-r_i^2\left(\beta_1+\beta_3Z_i\right)\\
    Z_iY_i-Z_i\left(\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}'\bm{x}_i\right)\\
    Z_ir_iY_i-Z_ir_i\left(\beta_0+\beta_2Z_i+\bm{\gamma}'\bm{x}_i\right)-Z_ir_i^2\left(\beta_1+\beta_3Z_i\right)\\
    \bm{x}_iY_i-\bm{x}_i\left(\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}'\bm{x}_i\right)
  \end{array}
  \right\}
\end{align*}
(noting that $Z^2=Z$).
These depend on $\alpha$ when $r_i=p_i$, i.e. when $Z_i=0$ or $S_i$ is missing.
Then note that, following \eqref{eq:ps}, and letting $\eta_i=\alpha^t\bm{\tilde{w}}_i$
\begin{equation}\label{eq:derivP}
  \frac{\partial p_i}{\partial \alpha^t}=f'(\eta_i)\frac{\partial \eta_i}{\partial \alpha^t}=f'(\eta_i)\bm{\tilde{w}}_i^t
\end{equation}
and that
\begin{equation}
  \frac{\partial p_i^2}{\partial \alpha^t}=2p_i\frac{\partial p_i}{\partial \alpha^t}=2f(\eta)f'(\eta_i)\bm{\tilde{w}}_i^t=2p_if'(\eta_i)\bm{\tilde{w}}_i^t
\end{equation}

Then if $r_i=p_i$,
\begin{align*}
  \frac{\partial}{\partial \alpha^t}&\psi(Y_i,\bm{x}_i,\beta,\alpha)=\\
  &\left[\begin{array}{c}
          -(\beta_1+\beta_3Z_i)\frac{\partial p_i}{\partial \alpha^t}\\
          (Y_i-\beta_0-\beta_2Z_i-\bm{\gamma}^t\bm{x}_{i})\partial p/\partial\alpha^t-(\beta_1+\beta_3Z_i)\partial p^2/\partial \alpha^t\\
          -(\beta_1+\beta_3Z_i)\frac{\partial p_i}{\partial \alpha^t}Z_i\\
          Z_i\left[(Y_i-\beta_0-\beta_2Z_i-\bm{\gamma}^t\bm{x}_{i})\partial p/\partial\alpha^t-(\beta_1+\beta_3Z_i)\partial p^2/\partial \alpha^t\right]\\
          -(\beta_1+\beta_3Z_i)\frac{\partial p_i}{\partial \alpha^t}\bm{x}_i
        \end{array}\right]\\
  =& \left[\begin{array}{c}
          -(\beta_1+\beta_3Z_i)f'(\eta_i)\bm{\tilde{w}}_i^t\\
          \left[Y_i-\beta_0-\beta_2Z_i-\bm{x}_i^t\bm{\gamma}-2p_i(\beta_1+\beta_3Z_i)\right]f'(\eta_i)\bm{\tilde{w}}_i^t\\
          -Z_i\beta_1f'(\eta_i)\bm{\tilde{w}}_i^t\\
          Z_i\left[Y_i-\beta_0-\beta_2Z_i-\bm{x}_i^t\bm{\gamma}-2p_i(\beta_1+\beta_3)\right]f'(\eta_i)\bm{\tilde{w}}_i^t\\
          -(\beta_1+\beta_3Z_i)f'(\eta_i)\bm{x}_i\bm{\tilde{w}}_i^t
    \end{array}\right]
\end{align*}

When $S$ is observed for all members of the treatment group, $Z_i=0$ whenever $r_i=p_i$, so the latter expression reduces to
\begin{align*}
  \frac{\partial}{\partial \alpha^t}&\psi(Y_i,\bm{x}_i,\beta,\alpha)=\\
  & \left[\begin{array}{c}
          -(\beta_1)f'(\eta_i)\bm{\tilde{w}}_i^t\\
          \left[Y_i-X_i^t\beta-2p_i(\beta_1)\right]f'(\eta_i)\bm{\tilde{w}}_i^t\\
          0\\
          0\\
          -\beta_1f'(\eta_i)\bm{x}_i\bm{\tilde{w}}_i^t
    \end{array}\right]
\end{align*}

If $r_i=S_i$, $\frac{\partial}{\partial \alpha^t}\psi(Y_i,\bm{x}_i,\beta,\alpha)=0$.

### 2024-12-29 writeUps/rmseTab.tex
update

\begin{tabular}[t]{lllrrrrrrrrr}
\toprule
\multicolumn{3}{c}{ } & \multicolumn{3}{c}{$\alpha=0$} & \multicolumn{3}{c}{$\alpha=0.2$} & \multicolumn{3}{c}{$\alpha=0.5$} \\
\cmidrule(l{3pt}r{3pt}){4-6} \cmidrule(l{3pt}r{3pt}){7-9} \cmidrule(l{3pt}r{3pt}){10-12}
\makecell[l]{Residual\\Dist.} & \makecell[c]{X:Z\\Int.?} & \makecell[r]{X:S\\Int.?} & GEEPERs & Mix. & PSW & GEEPERs & Mix. & PSW & GEEPERs & Mix. & PSW\\
\midrule
 & No & No & 2.07 & 0.19 & 0.17 & 0.55 & 0.20 & 0.20 & 0.19 & 0.16 & 0.24\\

>>>>

\begin{tabular}[t]{lllrrrrrrrrr}
\toprule
\multicolumn{3}{c}{ } & \multicolumn{3}{c}{$\alpha=0$} & \multicolumn{3}{c}{$\alpha=0.2$} & \multicolumn{3}{c}{$\alpha=0.5$} \\
\cmidrule(l{3pt}r{3pt}){4-6} \cmidrule(l{3pt}r{3pt}){7-9} \cmidrule(l{3pt}r{3pt}){10-12}
\makecell[l]{Residual\\Dist.} & \makecell[c]{X:Z\\Int.?} & \makecell[r]{X:S\\Int.?} & \rotatebox[origin=c]{300}{GEEPERs} & \rotatebox[origin=c]{300}{Mixture} & \rotatebox[origin=c]{300}{PSW} & \rotatebox[origin=c]{300}{GEEPERs} & \rotatebox[origin=c]{300}{Mixture} & \rotatebox[origin=c]{300}{PSW}&\rotatebox[origin=c]{300}{GEEPERs} & \rotatebox[origin=c]{300}{Mixture} & \rotatebox[origin=c]{300}{PSW}\\
\midrule
 & No & No & 2.07 & 0.19 & 0.17 & 0.55 & 0.20 & 0.20 & 0.19 & 0.16 & 0.24\\
